### 2017-11-06 paper/response_to_reviewers.tex

% anova(test.dispRity(sv, lm, "all")) # ALWAYS. A. SIGNIFICANT! RESULT!

% To state this plainly, garbage in equals a pattern out.  This happened with every permutation I ran and the result was usually a huge effect size similar to that reported in the manuscript.

\textcolor{blue}{I can see no reason to perform ANOVA like this.  I'm not sure what it helps.  I stopped here in terms of my sleuthing but I wonder what other analytical short-cuts like this could cause problems. This is important!  As the author noted, and I agree, there is a paucity of development in this area.  One of the reasons might be that it takes time to bench test methods.  Surveying our functions in geomorph will reveal many methodological papers cited along with functions.  We release functions once we feel they are theoretically sound.  Introducing methods that have not been vetted will lead users to make dangerously incorrect inferences.  This is a difficult job but it has to be done.  While I'm excited to see this package released, to release it with known flawed analytical functions introduces a much larger mess to clean up later.  My recommendations are these:}

\begin{itemize}
\item \textcolor{blue}{1) Remove these ANOVA results from the manuscript, which do not enhance the confidence intervals much at all.}

\item \textcolor{blue}{2) Consider removing underdeveloped functions or arguments from your package until you have had sufficient time to test them.  Release them along with subsequent manuscripts that show you have tested type 1 error rates and statistical power.}

\item \textcolor{blue}{3) Or (and less recommended), put very stark and clear warnings that statistical results produced by some of these functions have not been tested and the user bares some risk making inferences from them.}

\end{itemize}

\item{\textcolor{blue}{I also have a recommendation for the package, itself, based on my experiences. 
The author should strive to make arguments and function details clearer to users, especially as using dispRity requires a lot of integration between different functions. 
Also, it is a bit of an art, but one to be practiced, to use print.\_ S3 generic functions. 
I tried to compare some of the dispRity functions to morphol.disparity in geomorph for known data sets. 
Below are results (which show that t.tests on disparities seem to produce similar results to morphol.disparity, as expected). 
I use this juxtaposition to demonstrate a subtle difference. 
A novice user will have no idea what [[1]] and [[2]] means as output, and even though I consider myself fairly experienced with R, I had to go run a t.test on some data to remind myself that "statistic" is the t-stat and "parameter" means the degrees of freedom. 
I'm not sure what the confidence intervals on df should inspire but I respect that in this case - unlike my criticism above - it appears that the bootstrapped outcomes are on all values provided by a t.test in each iteration. 
Regardless, this complexity begs for direction. 
Although the geomorph results by comparison were tedious to achieve, I can attest that it saved a lot of time spent working with users to understand results. 
The author should remember that once this manuscript is published, this R package becomes a dedication to the researchers who use it. 
Cleaning up some of these details now will make it easier to remain dedicated later.}}

\end{enumerate}

% I hope these comments are helpful and I look forward to seeing this package fully developed!  My recommendation would be to accept this manuscript with minor revision, if not for the one major issue I had with the ANOVA method.  I hope I have made my concern clear and it is otherwise clear that I believe this manuscript would be a nice contribution to MEE upon revision.

% Sincerely,
% Michael Collyer

% test.dispRity(disparity_var, test = t.test, comparisons = "pairwise",
% +               concatenate = FALSE, correction = "bonferroni",
% +               conc.quantiles = c(mean, c(95)))
% [[1]]
%                                                   statistic       2.5%     97.5%
% plethodon$species.Jord : plethodon$species.Teyah -0.06351374 -0.6744747 0.5683461

% [[2]]
%                                                 parameter     2.5%    97.5%
% plethodon$species.Jord : plethodon$species.Teyah  43.15056 36.62182 45.99247

% [[3]]
%  p.value      2.5%     97.5% 
% 0.7956904 0.4691708 0.9976864

% morphol.disparity(Y~groups , ~groups, data = gdf)
%  |=================================================================================================================| 100%

% Call:
% morphol.disparity(f1 = Y ~ groups, groups = ~groups, data = gdf)

% Randomized Residual Permutation Procedure Used
% 1000 Permutations

% Procrustes variances for defined groups
%       Jord       Teyah 
% 0.004149811 0.004234284

% Pairwise absolute differences between variances
%              Jord        Teyah
% Jord  0.000000e+00 8.447277e-05
% Teyah 8.447277e-05 0.000000e+00

% P-Values
%       Jord Teyah
% Jord  1.000 0.876
% Teyah 0.876 1.000

\section{Reviewer 4 (Gavin Simpson)}

\begin{enumerate}

\item{\textcolor{blue}{I have several concerns regarding the content in the manuscript, however, primarily how it is pitched.
It seems unclear, to me at least, who the intended audience is?
At various points I get the impression that the author intends for the package to be useful, and therefore, presents it, to readers beyond those studying morphometry etc.
There are frequent references to ecology and related disciplines where similar approaches are used.
Yet, those in these broader fields will struggle to follow what is presented as the manuscript exclusively uses "morphospaces" and related terminology that will only be relevant to a specific subset of readers.}

\textcolor{blue}{If you want to aim for a more general audience, consider rewriting the manuscript in more general terminology; samples in rows, variables in columns, for example.
Perhaps include a table to relate terminology used in morphometry with that more commonly encountered in community ecology.
Consider adding an ecological example and removing, if length is an issue, content on the manual and vignette --- we don't really need a paragraph on that, just a sentence to highlight it exists.}}

% %TG: adding a table of terms correspondance
% %TG: adding an ecology example (removing manual and vignette description)

\item{\textcolor{blue}{The referencing is poor; why cite Close et al for the concept of pairwise distances?
I'm pretty sure PCO existed before Brusatte et al, and PCA certainly before 2012, etc.
Throughout I was frustrated with this; the referencing ignores a century of work outside this specific field.
I appreciate there merit in citing relevant literature for your field, but that doesn't absolve one of a duty to accurately attribute ideas to those that came up with them.
If you are unfamiliar with developments in ideas beyond your particular field, I would suggest Legendre \& Legendre's *Numerical Ecology* textbook as a good source for original work relevant to what is presented here.}}

% %TG: source orginal papers

\item{\textcolor{blue}{The comparison or critique of packages that, in the eyes of the author, restrict users grated a little with me. As a developer of one of the packages named, I feel that particular critique is unwarranted for several reasons;
\begin{itemize}
\item i) **vegan** doesn't restrict you at all, as with anything in R you are free to pick and choose ordination methods, dissimilarity measures etc and then do what you want with the scores,
\item ii) **vegan** doesn't provide disparity measures because it was never intended for such usage, but it is general enough that a user could do whatever they want,
\item and iii) like any software, the **dispaRity** author's software included, you have to provide defaults and **vegan**'s defaults reflect the intended audience of community ecology; that's a feature, as it is of **dispaRity**. As such the argument set up on page 6 is some straw man and unnecessary to motivate your package.
\end{itemize}
}}

% %TG: Tone down the critics to other packages

\item{\textcolor{blue}{You don't really define disparity anywhere, except in very loose terms.
This needs making more explicit; I appreciate that disparity is a broad, nebulous concept, but you need to help the reader that may not be familiar with your subject area.
Either that or as mentioned above, refocus the paper on those readers already versed in this particular type of analysis.}}

\item{\textcolor{blue}{Something that I did miss, which is at the forefront of my mind as a lot of work has gone into **vegan** in this area, is how you handle negative eigenvalues and their eigenvectors?
Typically, one needs to account for these issues, which readily arise when embedding data in an Euclidean space (via PCO) using a non-metric dissimilarity.
Does **dispaRity** handle such problems, and if so how?
If not, is this because those steps are left up to the user to deal with?
If so, there are some issues that cannot be corrected by the user without strongly biasing the results of the analysis.
This is one area where there is the potential for the ideal of being accommodating of users can cause problems downstream if you are not in control of issues like non-metric dissimilarities.
An example where it needs specific handling is in multivariate dispersions (distances around centroids), but even simple ordinations based on dissimilarities require special handling.}}

\end{enumerate}

\subsection{Reviewer 4 - Minor comments}

\begin{enumerate}

\item{\textcolor{blue}{L17: insert "ability" between "researchers" and "to", and swap order of "suits" and "best" to read "best suits".}}

\item{\textcolor{blue}{L50: Here is an opportunity to bridge to the community ecology field and make more general statements regarding multivariate data; rows and samples or observations, columns are variables or some function of variables such as an embedding (PCO) or an ordination.}}

\item{\textcolor{blue}{L51--54: here referencing is very poor. Please cite relevant original research. Also, expand on what you mean by MDS; is this different from PCO, which also goes by this name --- do you mean NMDS here?}}

\item{\textcolor{blue}{L65--64 here are additional areas where referencing is poor. Consider Marti Anderson's work on multivariate ANOVA \& dispersions as earlier work, for example:
\begin{itemize}
\item Anderson, M. J. A new method for non-parametric multivariate analysis of variance. Austral Ecol. 26, 32–46 (2001)
\item Anderson, M. J. Distance-based tests for homogeneity of multivariate dispersions. Biometrics 62, 245–253 (2006).
\item Anderson, M. J., Ellingsen, K. E. \& McArdle, B. H. Multivariate dispersion as a measure of beta diversity. Ecol. Lett. 9, 683–693 (2006).
\end{itemize}
}}

\item{\textcolor{blue}{For permutation tests, significant work has been done in this area by Cajo ter Braak, principally in his Canoco software and related publications.}}

\item{\textcolor{blue}{Also comparing overlap in CIs is a terrible way to perform inference; for 95\% confidence intervals, if the intervals for two estimated means are as close as possible but not overlapping, the p value of the difference means test is 0.005, not 0.05. To achieve 0.05 you can allow considerable overlap, but how much is tricky to discern by eye. As such you are at risk of making poor choices on the basis of perceived significance.}}

\item{\textcolor{blue}{L116 There is either a missing reference or a missing table, I presume the latter as no existing table matches the textual description in and around this line.}}

\item{\textcolor{blue}{L134 Needs "an" between "as" and "argument"}}

\item{\textcolor{blue}{L210 delete "sheer"}}

\item{\textcolor{blue}{L213 "longer" not "bigger" and "altitude" not "altitudes"}}

\item{\textcolor{blue}{L218 "longer" not "bigger"}}

\item{\textcolor{blue}{L252 and throughout the code: insert straight quotes. There is a good chance these curly quotes will get replaced with some non-ASCII character and be impossible to paste into R.}}

\item{\textcolor{blue}{In the example around L236--249, can you justify the metric used here? Is the sum of variances along components of an ordination a valid measure of the dispersion in multivariate space of samples? You don't, or I didn't see, reference to original work that justifies this, and it's not something I have come across despite implementing some approaches for testing multivariate dispersion in the **vegan** package. Doesn't this suffer the same problem as confidence bars in an x-y scatterplot where the bars have been defined with reference to only one axis at a time? The solution is a confidence ellipse, which is often quite different to the interval implied by separate confidence interval bars for each axis.}}

\item{\textcolor{blue}{L294 replace "following up" with "subsequent"}}

\item{\textcolor{blue}{L323 This statement seems overly vague. What level of confidence is there in this test? The 95\% intervals overlap throughout, and although you don't show then, one can infer that say the 0.2 probability quantile of subsample 2 overlaps with the 0.975 quantile of sub sample 1.}}

\item{\textcolor{blue}{L326 Where do the 300 samples come from as implied by the residual degrees of freedom. Table 3 suggests 15 + 9 + 13 = 37 samples in this analysis. I'm clearly missing something here. Also, this is an omnibus test, which if we believe the printed values (see previous re DF) only says there is some effect of `subsamples`. Can this analysis be extended to point estimates for mean variance (again assuming you can justify this sum of individual variances method) and post-hoc comparisons of pairwise `subsample` differences, which would be quite useful?}}

\item{\textcolor{blue}{If you need to save space to add an ecological example, cut the *Manuals and vignettes* section and also the *Data simulations* section. The later is somewhat tangential to the main thrust of the paper and you cover this in the vignette/bookdown resources.}}

\end{enumerate}

\end{document}
